---
output: html_document
keep_md : true
---

## Analyis on Weight Lifting Exercises done on HAR devices

### Executive Summary
The aim of this report is to use the data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants and predict the manner in which they did the exercise. More information is available from the website here (http://groupware.les.inf.puc-rio.br/har)

### Data Processing and Analysis
The training data for this project are available here:
[https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv](https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv).   
The test data are available here:
[https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv](https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv)

#### Load the required R packages

```{r message=FALSE}
library(caret)
library(randomForest)
## Load the training and test data sets
training_ds <- read.csv("pml-training.csv", header = TRUE, na.strings = c("NA", ""))
test_ds <- read.csv("pml-testing.csv", header = TRUE, na.strings = c("NA", ""))
dim(training_ds);dim(test_ds);
```
#### Handling Missing Values
```{r message=FALSE}
training_ds <- training_ds[, (colSums(is.na(training_ds)) == 0)]
test_ds <- test_ds[, (colSums(is.na(test_ds)) == 0)]
dim(training_ds);dim(test_ds);
```

#### Preparing for Training Set and Test Sets for cross validation
```{r message=FALSE}
inTrain = createDataPartition(y = training_ds$classe, p = 0.7, list = FALSE)
training = training_ds[inTrain, ]
cross_validation = training_ds[-inTrain, ]
dim(training);dim(cross_validation);

```
#### Remove some irrelvant fields to the prediction model in Training Set and Test Sets
```{r message=FALSE}
fields_to_delete <- grepl("X|user_name|raw_timestamp_part_1|raw_timestamp_part_2|cvtd_timestamp|new_window|num_window", colnames(training))
training <- training[, !fields_to_delete]
cross_validation <- cross_validation[, !fields_to_delete]
test_ds_final <- test_ds[, !fields_to_delete]
dim(training);dim(cross_validation);dim(test_ds_final)
```

#### Preprocessing with PCA by leaving the "classe" field
```{r message=FALSE}
preProc <- preProcess(training[, -length(training)], method = "pca", thresh = 0.99)
PCAtrain <- predict(preProc, training[, -length(training)])
PCAcross_validation <- predict(preProc, cross_validation[, -length(cross_validation)])
PCATest <- predict(preProc, test_ds_final[, -length(test_ds_final)])
dim(PCAtrain);dim(PCAcross_validation);dim(PCATest);
```

#### Train the prediction model with random forest algoritham and do cross validation testing
```{r message=FALSE}
ctrl <- trainControl(allowParallel=T,method = "cv", number = 4)
modelFit <- train(training$classe ~ ., method = "rf", data = PCAtrain,trControl = ctrl , importance = TRUE)

cross_predictions <- predict(modelFit, PCAcross_validation)
cMatrix <- confusionMatrix(cross_validation$classe, cross_predictions)
cMatrix$table
```
#### The output sample error
```{r message=FALSE}
accur <- postResample(cross_validation$classe, cross_predictions)
model_accuracy <- accur[[1]]
model_accuracy
```
The above model is `r model_accuracy` accurate.

#### Result Preditions with Teset Set
```{r message=FALSE}
predictions <- predict(modelFit, PCATest)

```
The prediction results from the test sets are as `r predictions`

